---
title: "functionalizing-rasters-from-NEON"
author: "Catherine Hulshof DLP"
date: "June 21, 2016"
output: html_document
---

```{r load-libraries }
library(raster)
library(rhdf5)
library(rgdal)
```

## Objectives

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)


```{r import raster }
# read LiDAR canopy height model
chm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")
```

1. For the CHM, set values == 0 to NA (not trees)

```{r assign-class }
# assign chm values of 0 to NA
chm[chm==0] <- NA

# do the values in the data look reasonable?
plot(chm,
     main="Canopy Height \n LowerTeakettle, California")
```

1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize density and plot vertical cutoff lines.

```{r histogram }

hist(chm,
     main="Distribution of Canopy Height  \n Lower Teakettle, California",
     xlab="Tree Height (m)",
     col="springgreen")
abline(v = 6,lwd=2,col="red")
abline(v=10,lwd=2,col="red")

```


```{r matrix-class }

# first create a matrix of values that represent the classification ranges

class.m <- c(0, 6, 1,
             6, 10, 2,
             10, 60, 3)
class.m

rcl.m <- matrix(class.m, 
                ncol=3, 
                byrow=TRUE)
rcl.m

# reclassify the raster using the reclass object - rcl.m
asp.ns <- reclassify(chm, 
                     rcl.m)

```

     
1. Plot the classified raster, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)


```{r plot }

# make room for a legend
par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 4.5))
# plot
plot(asp.ns,
     col=c("red","blue","green"), # hard code colors, unclassified (1)=red,
		 #N (2) =blue, S(3)=green
     main="Tree Canopy Height class  \n Lower Teakettle, California",
     legend=F)
# allow legend to plot outside of bounds
par(xpd=TRUE)
# create the legend
legend((par()$usr[2] + 20), 4103300,  # set x,y legend location
       legend = c("Grasses/Understorey", "Small Trees", "Tall Trees"),  # make sure the order matches the colors, next
       fill = c("red", "blue", "green"),
       bty="n") # turn off border
```

1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.

```{r export geotiff }

writeRaster(asp.ns,
            filename="../outputs/TEAK/Teak_Treeheightclass.tif",
            format="GTiff",
            options="COMPRESS=LZW",
            overwrite = TRUE,
            NAflag = -9999)

```


1. Export the plot figure to a pdf

```{r make pdf }

pdf("../outputs/TEAK/Teak_Treeheightclass.pdf", width=6, height=6)
# make room for a legend
par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 4.5))
# plot
plot(asp.ns,
     col=c("red","blue","green"), # hard code colors, unclassified (1)=red,
		 #N (2) =blue, S(3)=green
     main="Tree Canopy Height class  \n Lower Teakettle, California",
     legend=F)
# allow legend to plot outside of bounds
par(xpd=TRUE)
# create the legend
legend((par()$usr[2] + 20), 4103300,  # set x,y legend location
       legend = c("Grasses/Understorey", "Small Trees", "Tall Trees"),  # make sure the order matches the colors, next
       fill = c("red", "blue", "green"),
       bty="n") # turn off border

dev.off()
```




